# pre-class
###*Sarah Fringer*

####*September 17th, 2018*

```{r setup, include = FALSE, cache = FALSE}
library(knitr)
library(foreign)
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(results="hold")
knitr::opts_chunk$set(cache=F)
```

Make sure you commit this often with meaningfull messages. 

#### Background

The exponential distribution is defined by its cumulative distribution function
\(F(x) = 1-e^{-\lambda x}\)

The R function ***rexp()*** generates random variables with an exponential distribution. For example 
<center><strong>rexp(n=10, rate=5)</strong> </center>


results in 10 exponentially distributed numbers with a rate \(\lambda=5\). If you leave out the 5 and just have
<center><strong>rexp(n=10) </strong></center>
then this results in 10 exponentially distributed numbers with a rate \(\lambda=1\), this is also referred to as the "standard exponential distribution". 

### Part 1


1. Generate 200 random values from the standard exponential distribution and store them in a vector `exp.draws.1`.  Find the mean and standard deviation of `exp.draws.1`.

```{r}
exp.draws.1 <- rexp(n=200)
mean(exp.draws.1)
sd(exp.draws.1) 
```

2. Repeat, but change the rate to 0.2, 5, 7.3 and 10, storing the results in vectors called  `exp.draws.0.2`,  `exp.draws.5`,  `exp.draws.7.3` and  `exp.draws.10`. 

```{r}
exp.draws.0.2 <- rexp(n=200, rate=0.2)
exp.draws.5 <- rexp(n=200, rate=5)
exp.draws.7.3 <- rexp(n=200, rate=7.3)
exp.draws.10 <- rexp(n=200, rate=10)
mean(exp.draws.0.2)
sd(exp.draws.0.2) 
mean(exp.draws.5)
sd(exp.draws.5) 
mean(exp.draws.7.3)
sd(exp.draws.7.3) 
mean(exp.draws.10)
sd(exp.draws.10) 
```

3. The function `plot()` is the generic function in R for the visual display of data. `hist()` is a function that takes in and bins data as a side effect. To use this function, we must first specify what we'd like to plot.

a. Use the `hist()` function to produce a histogram of your standard exponential distribution. 

```{r}
hist(exp.draws.1)
```

b. Use `plot()` with this vector to display the random values from your standard distribution in order.
    
```{r}
plot(exp.draws.1)
```

c. Now, use `plot()` with two arguments -- any two of your other stored random value vectors -- to create a scatterplot of the two vectors against each other.
    
```{r}
plot(exp.draws.10, exp.draws.5)
```

4. We'd now like to compare the properties of each of our vectors. Begin by creating a vector of the means of each of our five distributions in the order we created them and saving this to a variable name of your choice. Using this and other similar vectors, create the following scatterplots and explain in words what is going on:
    a. The five means versus the five rates used to generate the distribution.
    
      * In this plot we can see that as the rate increases, the mean value of each distribution decreases. This would makes sense when looking at the values in our stored vectors for each distribution. For example, the values in the distribution with a rate of 10 are much lower than the values for the distribution with a rate of 0.2. 
      
    b. The standard deviations versus the rates.
    
      * In this plot we can see that as the rate increases, the standard                deviation value of each distribution decreases. This would makes sense when looking at the values in our stored vectors for each distribution. For example, we can see by looking at our stored vectors that the variance between each value in the distribution withrate 10 is much smaller than the variance between the values in the 0.2 rate distribution. 
      
    c. The means versus the standard deviations.
    
      * In this plot we can see a linear relationship between the mean and standard deviation of each distribution. This plot shows that the mean value and the standard deviation value are very similar for each distribution, providing a linear relationship. 

For each plot, explain in words what's going on.

```{r}
means <- c(1.022332, 4.36612, 0.2235881, 0.1447165, 0.08822352)
rates <- c(1, 0.2, 5, 7.3, 10)
sds <- c(1.098161, 4.739664, 0.2313931, 0.1431271, 0.0828559)
```

```{r}
library(ggplot2)
library(gridExtra)
library(stargazer)
plot1 <- qplot(means, rates)
plot2 <- qplot(sds, rates)
plot3 <- qplot(means, sds)
grid.arrange(plot1, plot2, plot3, ncol=3, nrow=1)
```

### Part II (PHP 2560 Only)


5. R's capacity for data and computation is large to what was available 10 years ago. 
  a. To show this, generate 1.1 million numbers from the standard exponential distribution and store them in a vector called `big.exp.draws.1`. Calculate the mean and standard deviation.

```{r}
big.exp.draws.1 <- rexp(n=1100000)
mean(big.exp.draws.1)
sd(big.exp.draws.1)
```
    
  b. Plot a histogram of `big.exp.draws.1`.  Does it match the function \(1-e^{-x}\)?  Should it? 
  
```{r}
hist(big.exp.draws.1)
```

 * Yes, this histogram matches the exponential function \(1-e^{-x}\). This makes sense because n is a very large number, 1.1 million, and as the sample size increases the histogram becomes more and more similar to the shape of the underlying/parent exponential distribution. 
 
  c. Find the mean of all of the entries in `big.exp.draws.1` which are strictly greater than 1. You may need to first create a new vector to identify which elements satisfy this.

```{r}
biggerthan1 <- big.exp.draws.1[big.exp.draws.1 > 1]
mean(biggerthan1)
```
  
  d. Create a matrix, `big.exp.draws.1.mat`, containing the the values in 
`big.exp.draws.1`, with 1100 rows and 1000 columns. Use this matrix as the input to the `hist()` function and save the result to a variable of your choice. What happens to your data?

```{r}
big.exp.draws.1.mat <- matrix(data = big.exp.draws.1, nrow = 1100, ncol = 1000, byrow=T)
histbigexp <- hist(big.exp.draws.1.mat)
```
  
  * The data looks very similar to the histogram made above when just plotting the vector of big.exp.draws.1. The only difference is only big.exp.draws.1.mat is a matrix, whereas big.exp.draws.1 is a vector. Not only do the two histograms look similar to each other, but they also both look simiilar to the histogram of the function \(1-e^{-x}\). 


  e. Calculate the mean of the 371st column of `big.exp.draws.1.mat`.
  
```{r}
meanofcol <- mean(big.exp.draws.1.mat[,371])
meanofcol
```
    
  f. Now, find the means of all 1000 columns of `big.exp.draws.1.mat` simultaneously. Plot the histogram of column means.  Explain why its shape does not match the histogram in problem 5b).

```{r}
meanofcols <- colMeans(big.exp.draws.1.mat)
hist(meanofcols)
```

* In problem 5b we were plotting the distribution of a sample exponential distribution. This shape followed the shape of the parent exponential distribtion function, which looks like a curve skewed to the right (Although the curve looks skewed, it is still normal because this is the normal shape of an exponential distribution). For this particular problem, 5f, we are instead plotting the sample means of the distribution. According to the central limit theorem as the sample size increases, its average value will be that of the standard normal N(0,1). In other words, since the mean is an unbiased estimator and is consistent, the sample mean will converge to the population mean of the parent exponential distribution and have a normal bell shaped curved histogram. 



